Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model

### 2014-03-25 Paper/gammaRays.tex
Fixed typo in table 2
Added reference to Castignani:2014gaa
Updated parameter values

Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics. Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} made possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}. These and previous studies led to several interesting results. For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}. This suggests that most of the bursts must have similar energetics in their rest frame. Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}. The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}. Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper). In this study we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.

@@

\begin{table}
	\centering
	\small
	\begin{tabular}{ l | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] }
		\multicolumn{1}{ c |}{GRB} & \multicolumn{2}{ c |}{$1\sigma$} & \multicolumn{2}{ c |}{$2\sigma$} & \multicolumn{2}{ c |}{$3\sigma$} & \multicolumn{2}{ c |}{$4\sigma$} & \multicolumn{2}{ c }{$5\sigma$} \\
		\hline
		080825C	&	1.04  & 2.24	&	0.67 & 3.32	&	0.42 & 5.83	&	0.19 & 14.9		&	0.087 & 35.7	\\
		090501	&	0.58  & 1.1		&	0.43 & 1.61	&	0.32 & 2.29	&	0.22 &  3.03	&	0.17  &  5.11	\\
		090902B	&	0.629 & 0.635	&	0.36 & 0.89	&	0.23 & 1.57	&	0.13 &  2.87	&	0.084 &  4.56	\\
		090926A &	2.61  & 3.33	&	1.99 & 6.62	&	1.34 & 9.15	&	0.73 & 13.5		&	0.48  & 19.4
	\end{tabular}
	\caption{Burst observation results. Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.}
	\label{tab:observationResults}
\end{table}

@@

@@

Finally, we are in a position to use the tests to fit parameters of the model:

@@

	}
	\item{$\eta_0$, which defines the luminosity scale.}
	\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
	\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
	\item{$\omega_0$, a characteristic radiation frequency;}
	\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
	\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
	\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
\end{itemize}

>>>>

Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics. Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} made possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}. These and previous studies led to several interesting results. For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}. This suggests that most of the bursts must have similar energetics in their rest frame. Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}. The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper). In this study we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.

@@

\begin{table}
	\centering
	\small
	\begin{tabular}{ l | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] }
		\multicolumn{1}{ c |}{GRB} & \multicolumn{2}{ c |}{$1\sigma$} & \multicolumn{2}{ c |}{$2\sigma$} & \multicolumn{2}{ c |}{$3\sigma$} & \multicolumn{2}{ c |}{$4\sigma$} & \multicolumn{2}{ c }{$5\sigma$} \\
		\hline
		080916C	&	1.04  & 2.24	&	0.67 & 3.32	&	0.42 & 5.83	&	0.19 & 14.9		&	0.087 & 35.7	\\
		090501	&	0.58  & 1.1		&	0.43 & 1.61	&	0.32 & 2.29	&	0.22 &  3.03	&	0.17  &  5.11	\\
		090902B	&	0.629 & 0.635	&	0.36 & 0.89	&	0.23 & 1.57	&	0.13 &  2.87	&	0.084 &  4.56	\\
		090926A &	2.61  & 3.33	&	1.99 & 6.62	&	1.34 & 9.15	&	0.73 & 13.5		&	0.48  & 19.4
	\end{tabular}
	\caption{Burst observation results. Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.}
	\label{tab:observationResults}
\end{table}

@@

Numerical values we used for these parameters are summarized in the table \ref{tab:parameters}.

\begin{table}
	\centering
	\small
	\begin{tabular}{ r | l | l }
		Parameter	&	Value			&	Comments \\
		\hline
		$\gamma$		&	$100$		&	\\
		$\eta_0$		&	$10^{45}$	&
	\end{tabular}
	\caption{Parameter values.}
	\label{tab:parameters}
\end{table}

@@

Finally, we can use the tests to fit parameters of the model:

@@

		Observations suggest values for relativistic factors of plasma is a range from $\gamma = 100$ to $\gamma = 1000$.
		We take $\gamma = 300$ as an order of magnitude estimate.
	}
	\item{
		$\eta_0$, which defines the luminosity scale.

		From photon counts.
	}
	\item{
		$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;

		From durations.
	}
	\item{
		$n$, which determines the sharpness of the jet end, $n > 3$;

	}
	\item{
		$\omega_0$, a characteristic radiation frequency;

		Non-physical
	}
	\item{
		$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;

	}
	\item{
		$k$, which determines how much the opening angle changes with frequency, $k < 0$;

	}
	\item{
		$\alpha$, the bare spectral index, $\alpha < -2k - 1$

	}
\end{itemize}
